TODO title

### 2017-01-16 doc/paper/paper.tex
Add references.

\section{Introduction}
Winner-take-all (WTA) mechanisms are often employed in cognitive models. For 
example, TODO O'Reilly, Nengo cleanup, TCM\@. A WTA mechanism receives 
a $d$-dimensional input of utility values for $d$ different choices. The output 
is supposed to be larger than zero for the dimension with highest utility and 
zero for all other inputs.

A large body of literature exists examining how well different WTA mechanisms 
can fit different data from decision experiments (TODO references). Here, 
however, we will investigate the suitability of two different WTA mechanisms in 
the context of large-scale cognitive modelling with spiking neurons on a set of 
benchmarks that are more normative in nature.  The first mechanism is an 
implementation of the classic Usher-McClelland (TODO ref) leaky, competing 
accumulator model.  It has been widely used, for example in versions of the TCM 
model (TODO ref), and TODO some of our models. We will compare this to an 
independent accumulator model with a second thresholding layer with recurrent 
connections to the first layer. In both cases we are especially interested in 
situations with more than two ($d > 2$) choices.

For the implementation of these models we use the Neural Engineering Framework 
(NEF) which allows us to directly implement the prescribed dynamics with spiking 
neurons. The NEF has been used in a wide range of models TODO, including the 
largest functional brain model Spaun (TODO ref). We will give a short 
introduction to the NEF first, then describe the implementation of the two WTA 
mechanisms. In section TODO, we give the results on a number of metrics, 
followed by a discussion in section TODO\@.

@@

\subsection{The Neural Engineering Framework}
We employ the Neural Engineering Framework (NEF, TODO cite) which allows us to 
implement mathematical model descriptions in spiking neural networks based on 
three principles: \emph{representation}, \emph{transformation}, and 
\emph{dynamics}.

@@

\subsection{Leaky, competing accumulator model}
We implement the widely used, leaky, competing accumulator model proposed by 
TODO\@. The dynamics for the state variables $x_i, i < N$, where $N$ is the 
number of choices, are given by
\begin{equation}\label{eqn:usher-mcclelland}
    \begin{split}
    dx_i &= \left(\rho_i - kx_i - \beta \sum_{j \neq i} x_j\right) 
    \frac{dt}{\tau'} \\
    x_i &\rightarrow \max(x_i, 0)
    \end{split}
\end{equation}
where $\rho_i$ is the external input, $k$ is the leak rate, $\beta$ the lateral 
inhibition, and $tau'$ the time constant. Note that one usually wants to set $k 
= \beta$ because TODO\@. Here, we chose $k = \beta = 1$. TODO what does changing 
this control? Why can we fix it at 1?

@@

\section{Results}
To test and compare the two WTA mechanisms we provide an input of $\rho_i 
= u - s \delta_{1i} + \eta_i$ where $u$ is the strength of the strongest input, 
$s$ the separation to all other inputs, $\delta$ the Kronecker delta, and 
$\eta_i$ Gaussian white noise with standard deviation $\sigma$. Thus, without 
loss of generality, the first state variable receives the strongest input and 
all other state variables receive input that is $s$ less. We chose this input as 
it it represents the worse case for picking out the strongest input as $s 
\rightarrow 0$ and as one input $\rho_i \rightarrow 0$ it will have less 
influence on the overall dynamics of the two WTA mechanisms.

@@

\section{Discussion}
We do not find either one of the networks to perform better on all metrics, but 
rather they excel at slightly different things. The Usher-McClelland network is 
better suited to quickly determine the correct winner. However, under noisy 
conditions it can end up not producing a clear output at all and basically not 
making a decision at all. The independent accumulator model might not be as 
quick or find the correct winner as often, but given enough time it will 
eventually arrive at a decision and produce a clear output.

@@

\nocite{ChalnickBillman1988a}
\nocite{Feigenbaum1963a}
\nocite{Hill1983a}
\nocite{OhlssonLangley1985a}
% \nocite{Lewis1978a}
\nocite{Matlock2001}
\nocite{NewellSimon1972a}
\nocite{ShragerLangley1990a}

@@

\bibliography{CogSci_Template}

>>>>

\section{Introduction}
Winner-take-all (WTA) mechanisms are often employed in cognitive 
models~\cite<e.g.,>{oreilly1998}.  A WTA mechanism receives a $d$-dimensional 
input of utility values for $d$ different choices. The output is supposed to be 
larger than zero for the dimension with highest utility and zero for all other 
inputs.

A large body of literature exists examining the optimality of WTA mechanisms and 
their consistency with neurobiological and psychological 
data~\cite<e.g.,>{bogacz2006,gold2007,smith2004}.  Here, however, we will 
investigate the suitability of two different WTA mechanisms in the context of 
large-scale cognitive modelling with spiking neurons on a set of benchmarks that 
are more normative in nature.  The first mechanism is an implementation of the 
classic leaky, competing accumulator model by~\citeA{usher2001}.  It has been 
widely used, for example in versions of the Temporal Context 
Model~\cite{sederberg2008}, and some of our models~\cite<e.g.,>{kajic2017}. We 
will compare this to an independent accumulator model with a second thresholding 
layer with recurrent connections to the first layer.  In both cases we are 
especially interested in situations with more than two ($d > 2$) choices.

For the implementation of these models we use the Neural Engineering 
Framework~\cite<NEF;>{eliasmith2003} which allows us to directly implement the 
prescribed dynamics with spiking neurons. The NEF has been used in a wide range 
of models, including the largest functional brain model 
Spaun~\cite{eliasmith2012}.  We will give a short introduction to the NEF first, 
then describe the implementation of the two WTA mechanisms. In 
Section~\ref{sec:results}, we give the results on a number of metrics, followed 
by a discussion in Section~\ref{sec:discussion}.

@@

\subsection{The Neural Engineering Framework}
We employ the Neural Engineering Framework~\cite<NEF;>{eliasmith2003} which 
allows us to implement mathematical model descriptions in spiking neural 
networks based on three principles: \emph{representation}, 
\emph{transformation}, and \emph{dynamics}.

@@

\subsection{Leaky, competing accumulator model}
We implement the widely used, leaky, competing accumulator model proposed 
by~\citeA{usher2001}.  The dynamics for the state variables $x_i, i < N$, where 
$N$ is the number of choices, are given by
\begin{equation}\label{eqn:usher-mcclelland}
    \begin{split}
    dx_i &= \left(\rho_i - kx_i - \beta \sum_{j \neq i} x_j\right) 
    \frac{dt}{\tau'} \\
    x_i &\rightarrow \max(x_i, 0)
    \end{split}
\end{equation}
where $\rho_i$ is the external input, $k$ is the leak rate, $\beta$ the lateral 
inhibition, and $tau'$ the time constant. Note that one usually wants to set $k 
= \beta$ because TODO\@. Here, we chose $k = \beta = 1$. TODO what does changing 
this control? Why can we fix it at 1?

@@

\section{Results}\label{sec:results}
To test and compare the two WTA mechanisms we provide an input of $\rho_i 
= u - s \delta_{1i} + \eta_i$ where $u$ is the strength of the strongest input, 
$s$ the separation to all other inputs, $\delta$ the Kronecker delta, and 
$\eta_i$ Gaussian white noise with standard deviation $\sigma$. Thus, without 
loss of generality, the first state variable receives the strongest input and 
all other state variables receive input that is $s$ less. We chose this input as 
it it represents the worse case for picking out the strongest input as $s 
\rightarrow 0$ and as one input $\rho_i \rightarrow 0$ it will have less 
influence on the overall dynamics of the two WTA mechanisms.

@@

\section{Discussion}\label{sec:discussion}
We do not find either one of the networks to perform better on all metrics, but 
rather they excel at slightly different things. The Usher-McClelland network is 
better suited to quickly determine the correct winner. However, under noisy 
conditions it can end up not producing a clear output at all and basically not 
making a decision at all. The independent accumulator model might not be as 
quick or find the correct winner as often, but given enough time it will 
eventually arrive at a decision and produce a clear output.

@@

@@

\bibliography{references}
